Data preparation

### 2023-09-27 procesamiento/prod_analysis-robust.Rmd
cloud

---
title: "Robustness checks"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "input/css/custom.css" # custom CSS para html
linestretch: '1.5'          # interlineado 
link-citations: yes         # citas linkeadas
output:
  bookdown::html_document2:
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: false    
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    pandoc_args:
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion  
---

@@

dfreg <- df1 %>% dplyr::select(redist=redistF,
                        class3,
                        homclass,
                        Q05pcm,
                        Q03pcm,
                        gini_disp,
                        gini_mkt,
                        giniot=gini,
                        rgdpna,
                        edyears,
                        female,
                        agenum,
                        country2,country,
                        union,
                        labst,
                        labst1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         age2=agenum^2,
         edyears2=edyears^2) %>%
  # filter(!country2 %in% c("Turkey","Estonia")) %>%
  na.omit()
# dim(dfreg)

>>>>

---
title: "Robustness checks"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css" # custom CSS para html
linestretch: '1.5'          # interlineado 
link-citations: yes         # citas linkeadas
output:
  bookdown::html_document2:
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: false    
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    pandoc_args:
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion  
---

@@

dfreg <- df1 %>% dplyr::select(redist=redistF,
                        class3,class5,
                        homclass,
                        Q05pcm,
                        Q03pcm,
                        # isei08,
                        gini_disp,
                        gini_mkt,
                        giniot=gini,
                        giniindex,
                        rgdpna,
                        edyears,
                        female,
                        agenum,
                        country2,country,
                        union,
                        labst,
                        labst1,
                        isolation,
                        MAINSTAT,
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         age2=agenum^2,
         edyears2=edyears^2) %>%
  # filter(!country2 %in% c("CHN")) %>%
  # filter(!country2 %in% c("TWN")) %>%
  # filter(!country2 %in% c("ZAF")) %>%
  # filter(!country2 %in% c("FIN")) %>%
  # filter(!country2 %in% c("USA")) %>%
  # filter(!country2 %in% c("GBR")) %>%
  # filter(!country2 %in% c("ISL")) %>%
  # filter(!country2 %in% c("TUR","EST")) %>%
  # filter(MAINSTAT %in% c(1,2)) %>%
  # filter(agenum %in% c(25:65)) %>% # Paskov & Weisstanner
  na.omit()

# lineal
base <- clmm(redist~1 + (1|country2),data=dfreg); icc1<- performance::icc(base)
class <- clmm(redist~ class3+ (1|country2),data=dfreg)
homclass <- clmm(redist~homclass + (1|country2),data=dfreg)
full1 <- clmm(redist~ class3 +homclass+ female+agenum + age2+(1|country2),data=dfreg)
# rob1 <- clmm(redist~class3+ homclass+ edyears+ female+ agenum+ age2+(1|country2),data=dfreg)
# rob2 <- clmm(redist~class3+ homclass+ edyears+ Q03pcm+ female+ agenum +age2+(1|country2),data=dfreg)
rob3 <- clmm(redist~class3+ homclass+ edyears+ Q03pcm+ labst1+ union+ female+agenum +age2+(1|country2),data=dfreg)
# rob4 <- clmm(redist~class3+ homclass+ edyears+ Q03pcm+ labst1+ union+ female+agenum +age2+(homclass|country2),data=dfreg)
# anova(rob3,rob4)
# Interactions segregation x class
int_homo <- clmm(redist~class3*homclass+ edyears+ Q03pcm+ labst1+ union+ female+agenum +age2+ (1|country2),data=dfreg)
# models <- list(homclass,class,full1,
#                # rob1,rob2,
#                rob4)
# texreg::screenreg(int_homo,single.row = F)

### 2023-09-27 procesamiento/prod_prep-macro.Rmd
cloud

# Economic inequality

```{r world-bank, eval=FALSE, include=FALSE}
# install.packages("wbstats")
library(wbstats)
ineq<- c("3.0.Rate75-25","3.0.Rate90-10","SI.POV.GINI","3.0.Gini","3.0.Gini_nozero")
# Gini
giniwb <- wb_data("SI.POV.GINI", start_date = 2015, end_date = 2020,return_wide = F)
gini_wb<- reshape::cast(giniwb[,c("iso3c","date","value")],formula = iso3c ~ date,value = "value")
gini_wb$ccode <- countrycode::countrycode(sourcevar = gini_wb$iso3c,origin = "iso3c",destination = "iso3n")
gini_wb<- gini_wb %>% filter(ccode %in% countries$ccode)

skimr::skim(gini_wb)

gini_wb$giniwb <-  gini_wb$"2017"
summary(gini_wb$giniwb)
gini_wb$giniwb <- ifelse(is.na(gini_wb$giniwb),yes = gini_wb$"2018",no = gini_wb$giniwb)
summary(gini_wb$giniwb)
gini_wb$giniwb <- ifelse(is.na(gini_wb$giniwb),yes = gini_wb$"2016",no = gini_wb$giniwb)
summary(gini_wb$giniwb)

# Rate 90-1
rate90_10 <- wb_data("3.0.Rate90-10", start_date = 2015, end_date = 2020,return_wide = F)
rate90_10_wb<- reshape::cast(giniwb[,c("iso3c","date","value")],formula = iso3c ~ date,value = "value")
rate90_10_wb$ccode <- countrycode::countrycode(sourcevar = rate90_10_wb$iso3c,origin = "iso3c",destination = "iso3n")
rate90_10_wb<- rate90_10_wb %>% filter(ccode %in% country_data$ccode)

# Rate 90-1
rate75_25 <- wb_data("3.0.Rate75-25", start_date = 2015, end_date = 2020,return_wide = F)
rate75_25_wb<- reshape::cast(giniwb[,c("iso3c","date","value")],formula = iso3c ~ date,value = "value")
rate75_25_wb$ccode <- countrycode::countrycode(sourcevar = rate75_25_wb$iso3c,origin = "iso3c",destination = "iso3n")
rate75_25_wb<- rate75_25_wb %>% filter(ccode %in% country_data$ccode)

country_data <- 
countries %>% left_join(select(gini_wb,iso3c,ccode,giniwb),by="ccode")
```

```{r WIID}
wiid<- haven::read_dta(here::here("input/data/original/WIID_30JUN2022_0/WIID_30JUN2022_0.dta"))
summary(wiid$gini)

gini_wiid <- wiid %>% dplyr::select(c3,year,gini,palma) %>% dplyr::filter(year %in% c(2014:2019))
summary(gini_wiid)

gini_wiid <- panelr::widen_panel(data = panelr::panel_data(as.data.frame(gini_wiid), id = c3 , wave = year))
gini_wiid$ccode <- countrycode::countrycode(sourcevar = gini_wiid$c3,origin = "iso3c",destination = "iso3n")

gini_wiid<- gini_wiid %>% filter(ccode %in% countries$ccode)

gini_wiid$giniwd <- gini_wiid$gini_2017
summary(gini_wiid$giniwd)
gini_wiid$giniwd <- ifelse(is.na(gini_wiid$giniwd),gini_wiid$gini_2018,no = gini_wiid$giniwd)
summary(gini_wiid$giniwd)
gini_wiid$giniwd <- ifelse(is.na(gini_wiid$giniwd),gini_wiid$gini_2016,no = gini_wiid$giniwd)
summary(gini_wiid$giniwd)

skimr::skim(gini_wiid)

gini_wiid$palmawd <- gini_wiid$palma_2017
summary(gini_wiid$palmawd)
gini_wiid$palmawd <- ifelse(is.na(gini_wiid$palmawd),gini_wiid$palma_2018,no = gini_wiid$palmawd)
summary(gini_wiid$palmawd)
gini_wiid$palmawd <- ifelse(is.na(gini_wiid$palmawd),gini_wiid$palma_2016,no = gini_wiid$palmawd)
summary(gini_wiid$palmawd)

country_data2 <- 
countries %>% left_join(gini_wiid[,c("ccode","c3","giniwd","palmawd")],by="ccode")
```

```{r Solt}
load(file = here::here("input/data/original/swiid9_5/swiid9_5.rda"))
solt <- 
swiid_summary %>% dplyr::select(country,year,gini_disp,gini_mkt) %>% dplyr::filter(year %in% c(2016))
solt$ccode <- countrycode::countrycode(sourcevar = solt$country,origin = "country.name",destination = "iso3n")
solt$country <- NULL
solt <-  solt %>% filter(ccode %in% country_data2$ccode)
country_data3 <- country_data2 %>% 
  left_join(solt,by="ccode")
```

```{r wid-dataset}
wid <- 
  read.csv(file = here::here("input/data/original/wid-percentiles/WID_Data_Metadata/WID_Data_28072023-104940.csv"),sep = ";")
country_data3 <- country_data3 %>% 
  left_join(wid,by="ccode")
summary(country_data3)
```

```{r otero23}
otero <- 
  read.csv(file = here::here("input/data/original/otero23.csv"),sep = ";")
otero$ccode <- countrycode::countrycode(sourcevar = otero$iso3n,origin = "iso2c",destination = "iso3n")
country_data3 <- country_data3 %>% 
  left_join(otero,by="ccode")
summary(country_data3)
```

# GDP

```{r pen-data}
pendata<- haven::read_dta(file = here::here("input/data/original/pen-world/pwt1001.dta"))
pendata$ccode <- countrycode::countrycode(sourcevar = pendata$countrycode,origin = "iso3c",destination = "iso3n")
pen <- pendata %>% filter(ccode %in% countries$ccode,year %in% 2017)

skimr::skim(pen)
country_data4 <- country_data3 %>% 
  left_join(pen,by="ccode")
```

```{r merge}
df1$ccode <- as.numeric(df1$country)
df2 <- 
df1 %>% left_join(country_data4[,c("ccode","c3","giniwd","gini_disp","gini_mkt","gini","gv_spen","palmawd","rgdpna","pc10_2017","pc50_2017")],by = "ccode")

df2$country2 <- countrycode::countrycode(sourcevar = df2$country,origin = "iso3n",destination = "country.name")
df2 <- df2 %>% rename(country3=c3)
df2 <- df2 %>% select(country,country2, country3, everything())
```

>>>>

**World Income Inequality Database (WIID)**

> UNU-WIDER, World Income Inequality Database (WIID) Companion dataset (wiidcountry and/or wiidglobal). Version 30 June 2022. https://doi.org/10.35188/UNU-WIDER/WIIDcomp-300622

```{r , eval=FALSE, include=FALSE}
wiid_long<- haven::read_dta(here::here("input/data/original/WIID_30JUN2022_0/WIID_30JUN2022_0.dta"))
wiid_long$ccode <- countrycode::countrycode(sourcevar = wiid_long$c3,origin = "iso3c",destination = "iso3n")

wiid_long1 <- 
wiid_long %>% 
  filter(ccode %in% countries$ccode) %>% 
  filter(year==2017)

```

```{r}
wiid <- 
  read.csv(file = here::here("input/data/original/wid-percentiles/wiid-data-2015-2020.csv"),sep = ";")
names(wiid) <- tolower(names(wiid))
names(wiid) <- gsub('\\.', '', names(wiid))
wiid$ccode <- countrycode::countrycode(sourcevar = wiid$iso,origin = "iso3c",destination = "iso3n")
wiid_2017 <- 
wiid %>% 
  filter(ccode %in% countries$ccode) %>% 
  filter(year==2017) %>% 
  select(ccode,everything(),-iso,-country,-year) 
```

**Solt SWDIID**
```{r Solt}
load(file = here::here("input/data/original/swiid9_5/swiid9_5.rda"))
solt <- 
swiid_summary %>% dplyr::select(country,year,gini_disp,gini_mkt) %>% dplyr::filter(year %in% c(2016))
solt$ccode <- countrycode::countrycode(sourcevar = solt$country,origin = "country.name",destination = "iso3n")
solt$country <- NULL
solt_2017 <-  solt %>% filter(ccode %in% countries$ccode) %>% 
  select(ccode,gini_disp,gini_mkt)
```

**Otero et.al (2023)**

```{r otero23}
otero <- 
  read.csv(file = here::here("input/data/original/otero23.csv"),sep = ";")
otero$ccode <- countrycode::countrycode(sourcevar = otero$iso3n,origin = "iso2c",destination = "iso3n")
otero$gdp <- gsub('\\,', '',otero$gdp) 
otero<- otero %>% select(ccode,gv_spen,individualism,gini,gdp)
```

**GDP: PEN data**
```{r pen-data}
pendata<- haven::read_dta(file = here::here("input/data/original/pen-world/pwt1001.dta"))
pendata$ccode <- countrycode::countrycode(sourcevar = pendata$countrycode,origin = "iso3c",destination = "iso3n")
pen_2017 <- pendata %>% filter(ccode %in% countries$ccode,year %in% 2017) %>% select(ccode,everything(),-country,-countrycode,-currency_unit,-year)
```

```{r merge}
macro_data <- 
  left_join(wiid_2017,solt_2017) %>%
  left_join(pen_2017) %>% 
  left_join(otero)

df1$ccode <- as.numeric(df1$country)
df2 <- left_join(df1,macro_data,by="ccode")
df2 <- 
df2 %>% mutate(country2=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "iso3c"),
               country3=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "country.name")) %>% 
  select(ccode,country2,country3,everything())
```
